---
format: 
  pdf:
    classoption: [landscape]

editor: visual
---

Title: Connectivity of stormwater ponds impacts Odonata abundance and species richness

Journal: Landscape Ecology

Authors: Richmond, Isabella C.\*, Perron, Mary Ann C., Boyle, Sean B., & Pick, Frances R.

\* Department of Biology, 30 Marie Curie Private, University of Ottawa, Ottawa, Ontario K1N 6N5, Canada Department of Biology, 7141 Sherbrooke St. W., Concordia University, Montreal, QC, Canada, H4B 1R6

email: [isabella.richmond\@mail.concordia.ca](isabella.richmond@mail.concordia.ca), phone : 438-439-8064

**Supplementary Information**

This Supplementary Information provides fine-scale details for each of our study ponds. Table S1-1 outlines descriptive characteristics of each pond, like age and area. Table S1-2 details the observed odonate data at each pond, including abundance, species richness, and shannon diversity. Table S1-3 describes the connectivity data at each pond, including mean and standard variation for current values, as well as number of nearest neighbours. All code and data for this project can be found at https://github.com/icrichmond/OdonataConnectivity.

\newpage

Table S1-1. Pond descriptive characteristics. SWF = stormwater pond, NAT = natural pond. Coordinates are in WGS 84. Some data were not available for each pond, indicated by NA.

```{=tex}
\begingroup
\setlength{\LTleft}{0pt minus 500pt}
\setlength{\LTright}{0pt minus 500pt}
\fontsize{11pt}{11pt}\selectfont
\addtolength{\tabcolsep}{-3pt}
```
```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(gt)

setwd("C:/Users/I_RICHMO/Documents/Repositories/odonata-connectivity")

# data
ani <- read.csv('input/cleaned/AnisopteraCleaned.csv', na.strings = "na") %>%
  mutate(DragonflyAbundance = abundance,
         DragonflySpeciesRich = speciescount, 
         DragonflyShannon = shannon) %>%
  select(Pond, Group, DragonflyAbundance, DragonflySpeciesRich, DragonflyShannon, mean.nine, sd.nine)
zyg <- read.csv('input/cleaned/ZygopteraCleaned.csv', na.strings = "na")%>%
  mutate(DamselflyAbundance = abundance,
         DamselflySpeciesRich = speciescount, 
         DamselflyShannon = shannon) %>%
  select(Pond, DamselflyAbundance, DamselflySpeciesRich, DamselflyShannon, mean.three, sd.three)
hab <- read.csv('input/cleaned/HabitatNumbers.csv', na.strings = "na") %>%
  select(-X)
ponds <- read.csv('input/pond_descriptive.csv', na.strings = "na")


# join
anizyg <- full_join(ani, zyg, by = "Pond")
pondshab <- left_join(hab, ponds, by = "Pond")
all <- full_join(anizyg, pondshab, by = "Pond")

# clean up
df<- all %>%
  select(Pond, Group, X, Y, age..years., area..m2., catchment..km2., imperviousness.of.catchment....) %>%
  rename('X coordinate' = X,
         'Y coordinate' = Y,
         'Age (years)' = age..years.,
         'Area (m2)' = area..m2., 
         'Catchment Area (km2)' = catchment..km2.,
         'Imperviousness of Catchment (%)' = imperviousness.of.catchment....) %>%
  gt() %>%
  as_latex()

df
```

```{=tex}
\endgroup
```
\newpage

```{=tex}
\begingroup
\setlength{\LTleft}{0pt minus 500pt}
\setlength{\LTright}{0pt minus 500pt}
\fontsize{7pt}{9pt}\selectfont
\addtolength{\tabcolsep}{-3pt}
```
Table S1-2. Pond odonate data. All sampling was done on adult odonates. SWF = stormwater pond, NAT = natural pond.

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

df<- all %>%
  select(Pond, Group, DragonflyAbundance, DragonflySpeciesRich, DragonflyShannon, DamselflyAbundance, DamselflySpeciesRich, DamselflyShannon) %>%
  rename('Dragonfly Abundance' = DragonflyAbundance,
         'Dragonfly Species Richness' = DragonflySpeciesRich, 
         'Dragonfly Shannon Diversity' = DragonflyShannon,
         'Damselfly Abundance' = DamselflyAbundance,
         'Damselfly Species Richness' = DamselflySpeciesRich,
         'Damselfly Shannon Diversity' = DamselflyShannon) %>%
  gt()

df
```

```{=tex}
\endgroup
```
\newpage

```{=tex}
\begingroup
\setlength{\LTleft}{0pt minus 500pt}
\setlength{\LTright}{0pt minus 500pt}
\fontsize{8pt}{9pt}\selectfont
\addtolength{\tabcolsep}{-3pt}
```
Table S1-3. Pond connectivity data. SWF = stormwater pond, NAT = natural pond. 900 m scale was used for dragonflies, 300 m scale was used for damselflies. Higher values of mean current indicate higher connectivity. Current was calculated using *Circuitscape*.

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

df<- all %>%
  select(Pond, Group, mean.nine, n.nine, mean.three, n.three) %>%
  rename('Mean Current Density (900 m)' = mean.nine,
         'Mean Current Density (300 m)' = mean.three,
         'Number of Nearest Neighbours (900 m)' = n.nine,
         'Number of Nearest Neighbours (300 m)' = n.three) %>%
  gt() %>%
  as_latex()

df
```

```{=tex}
\endgroup
```
